\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-09-15 submission/manuscript.Rmd
intro

<!-- Attractiveness -->
With the advancement of technology, AI-generated images of faces have increasingly been used to study various cognitive mechanisms and processes, such as that underlying emotion processing [@dyck2008recognition; @dyck2010virtual]. Notably, the ability to manipulate the degree of emotionality and salience of virtual faces, by changing its facial dimensions, make them ideal targets for simulation monitoring research [@fuentes2018evolutionary; @ferstl2018perceptual]. To this effect, the attractiveness of AI-generated faces, which integrates both components of emotional intensity and saliency [@o2003beauty; @debruine2007dissociating; sobieraj2014beautiful], have indeed been associated with its perceived realness. To be specific, research findings suggest that artificially created faces were judged to be less attractive as compared to real faces [@liefooghe2022faces; tsikandilakis2019beauty; @diel2022familiarity]. Interestingly, @liefooghe2022faces further reports that attractiveness ratings differed significantly between participants who were told that the faces were AI-generated from those who had no prior knowledge. Whereas such studies posit that lower perceived realness of a given stimulus leads to lower ratings of attractiveness, little is known about the impacts of attractiveness on simulation monitoring.

<!-- Goals and hypotheses -->
The goal of this study is therefore to determine whether the attractiveness of artificially-generated faces affects our reality judgement of it. Based on the Embodied Reality Theory [@makowski2018cognitive; @makowski2019phenomenal], which suggests that highly salient stimuli are perceived to be more real, we hypothesize a quadratic relationship between the perceived realness and attractiveness of a given virtual stimulus, i.e., faces rated as highly attractive or unattractive will be judged to be more real. Additionally, we further postulate that inter-individual characteristics, such as personality traits and predisposed attitudes, will interact with the relationship between perceived realness and attractiveness. Moreover, given the well-established positive correlation between attractiveness and perceived trustworthiness as well as familiarity often reported in previous studies, we posit that virtual faces judged to be real will also be rated as more trustworthy and familiar [@liefooghe2022faces; @bartosik2021you; @garrido2017kdef; @little2011facial].

>>>>

<!-- Attractiveness -->
AI-generated images of faces, due to their popularity as a target of CGI technology and the possibility of experimentally manipulating facial features, are increasingly used to study face processing as related to saliency or emotions, as well as to other important components of faces evaluation, such as trustworthiness or attractiveness (**REF**). Interestingly, several studies report that artificially created faces were judged to be less attractive (**QUESTION: was their nature known to the participant?**) as compared to real faces [@tsikandilakis2019beauty; @diel2022familiarity], and @liefooghe2022faces further reports that attractiveness ratings differed significantly between participants who were told that the faces were AI-generated from those who had no prior knowledge. Whereas this line of evidence suggests that beliefs of reality have an effect on attractiveness ratings of faces, the opposite question, whether attractiveness could drive beliefs of reality, has received no attention to date.

<!-- Goals and hypotheses -->
This study aims primarily at exploring the effect of face attractiveness on simulation monitoring, i.e., on the beliefs that an image is real or artificially generated. Based on the embodied reality theory [outlined in @makowski2018cognitive; @makowski2019phenomenal], which suggests that salient and emotional stimuli are perceived to be more real, we hypothesize a quadratic relationship between the perceived realness and attractiveness: faces rated as highly attractive or unattractive will more likely believed to be real. Moreover, we expect a similar relationship with trustworthiness ratings given its well-established link with attractiveness [@liefooghe2022faces; @bartosik2021you; @garrido2017kdef; @little2011facial], and a positive relationship with familiarity (as more familiar faces would appear as more salient, self-relevant and anchored in reality). Additionally, we will further explore the role of dispositional traits, such as personality and attitude towards AI, characteristics on inter-individual simulation monitoring tendencies.
